---
title: "Acute COPD Exacerbation Prediction Tool (ACCEPT)"
site: "bookdown::bookdown_site"
runtime: "shiny"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
    delete_merged_file: TRUE
    table_css: FALSE
    output_dir: "docs"
    css: ["css/pretty-styles.css", "css/cayman.css", "css/python.css", "css/shadow.css", "index.html"]
    js: ["js/list-group.js", "js/code-chunk.js", "js/code-chunk-component.js", "js/highlight.js"]
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: null
      search: yes
      fontsettings:
        theme: white
        family: sans
        size: 2
      sharing:
        facebook: yes
        twitter: yes
        google: no
        linkedin: no
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'google', 'twitter', 'linkedin', 'weibo', 'instapaper']
  bookdown::pdf_book:
    keep_tex: yes
---
# The Study {#thestudy}
The **accept** package and application are based on the paper [Development and External Validation of the Acute COPD Exacerbation Prediction Tool (ACCEPT)](https://www.biorxiv.org/content/biorxiv/early/2019/06/04/651901.full.pdf), published in 2019.

## Background {#background}

Chronic Obstructive Pulmonary Disease (COPD) affects 300 million people worldwide. Patients with COPD often experience what is called an **acute exacerbation**.

**exacerbation**: acute worsening of COPD symptoms; acute means sudden onset, short in duration; COPD symptoms include shortness of breath, wheezing, coughing up mucus, etc.

### Exacerbation Causes {-}

```{r, eval=TRUE, echo=FALSE}
source("R/addHTML.R")
addHTML("exacerbation-cause.html", section = 2, codeChunk = FALSE)
```

### Exacerbation Severity {-}

```{r, eval=TRUE, echo=FALSE}
addHTML("exacerbation-severity.html", section = 2, codeChunk = FALSE)
```

### References {-}

[COPD exacerbations: definitions and classifications](https://erj.ersjournals.com/content/erj/21/41_suppl/46s.full.pdf)

[Toward a consensus definition for COPD exacerbations](https://www.ncbi.nlm.nih.gov/pubmed/10843984)

## The Study Population {#studyPopulation}

The data for this study was taken from three different studies:

[MACRO Clinical Trial](https://clinicaltrials.gov/ct2/show/NCT00325897)

[MACRO Paper](https://www.ncbi.nlm.nih.gov/pubmed/21864166)

[STATCOPE Clinical Trial](https://clinicaltrials.gov/ct2/show/NCT01061671)

[STATCOPE Paper](https://www.ncbi.nlm.nih.gov/pubmed/24836125?dopt=Abstract)

[OPTIMAL Clinical Trial](https://www.ncbi.nlm.nih.gov/pubmed/15611808)

[OPTIMAL Paper](https://annals.org/aim/fullarticle/734106)

### Clinical Trial 1: MACRO {-}

Number of Participants: 1142

Type of Study: Clinical Randomized

Masking: Triple (Participant, Investigator, Care-Provider)

#### Criteria {-}

- Age: > 40 years old
- COPD diagnosis
- Does not have asthma
- 10 pack years or more of smoking

#### Study Groups {-}

```{r diagram, eval = TRUE, echo = FALSE}
library(DiagrammeR)
nodes <- create_node_df(n=3, type = "character",
                        color = themeColors[1],
                        fontcolor = "white",
                        fillcolor = themeColors[1],
                        shape = "oval",
                        fontsize = 4,
                        label = c("All \n Participants","Placebo",
                                  "Azithromycin \n Treatment"))

edges <- create_edge_df(from = c(1, 1),
                        to = c(2, 3),
                       rel = "leading to",
                       color = themeColors[2])

graph <- create_graph(nodes_df = nodes,
                      edges_df = edges)
render_graph(graph, layout="tree")
```


```{r, eval=TRUE, echo=FALSE}
addHTML("study1-arms.html", section = 2, codeChunk = FALSE)
```

#### Data Collection: Primary Metric {-}

```{r, eval=TRUE, echo=FALSE}
addHTML("study1-primary-metric.html", section = 2, codeChunk = FALSE)
```

### Clinical Trial 2: STATCOPE {-}

Number of Participants: 885

Type of Study: Clinical Randomized

Masking: Triple (Participant, Investigator, Care-Provider)

#### Criteria {-}

- Age: 40 - 80 years old
- COPD diagnosis
- Does not have asthma
- 10 pack years or more of smoking

#### Study Groups {-}

```{r diagram2, eval = TRUE, echo = FALSE}
library(DiagrammeR)
nodes <- create_node_df(n=3, type = "character",
                        color = themeColors[1],
                        fontcolor = "white",
                        fillcolor = themeColors[1],
                        shape = "oval",
                        fontsize = 4,
                        label = c("All \n Participants","Placebo",
                                  "Simvastatin \n Treatment"))

edges <- create_edge_df(from = c(1, 1),
                        to = c(2, 3),
                       rel = "leading to",
                       color = themeColors[2])

graph <- create_graph(nodes_df = nodes,
                      edges_df = edges)
render_graph(graph, layout="tree")
```


```{r, eval=TRUE, echo=FALSE}
addHTML("study2-arms.html", section = 2, codeChunk = FALSE)
```

#### Data Collection: Primary Metric {-}

```{r, eval=TRUE, echo=FALSE}
addHTML("study2-primary-metric.html", section = 2, codeChunk = FALSE)
```

### Clinical Trial 3: OPTIMAL {-}

Number of Participants: 432

Type of Study: Clinical Randomized

Masking: Double (Participant, Investigator)

#### Criteria {-}

- Age: > 35 years old
- COPD diagnosis
- Does not have asthma
- 10 pack years or more of smoking
- at least 1 COPD exacerbation in the last year

#### Study Groups {-}

```{r diagram3, eval = TRUE, echo = FALSE}
library(DiagrammeR)
nodes <- create_node_df(n=4, type = "character",
                        color = themeColors[1],
                        fontcolor = "white",
                        fillcolor = themeColors[1],
                        shape = "oval",
                        fontsize = 4,
                        label = c("All \n Participants","Placebo \n Inhaler",
                                  "Salmeteral \n Inhaler",
                                  "Advair \n Inhaler"))

edges <- create_edge_df(from = c(1, 1, 1),
                        to = c(2, 3, 4),
                       rel = "leading to",
                       color = themeColors[2])

graph <- create_graph(nodes_df = nodes,
                      edges_df = edges)
render_graph(graph, layout="tree")
```


```{r, eval=TRUE, echo=FALSE}
addHTML("study3-arms.html", section = 2, codeChunk = FALSE)
```

#### Data Collection: Primary Metric {-}

```{r, eval=TRUE, echo=FALSE}
addHTML("study3-primary-metric.html", section = 2, codeChunk = FALSE)
```

## Data Collection: Combined Trials {#combined-trials}

For our study, we combined the results from the three trials described above.

### Primary Metric {-}

```{r, eval=TRUE, echo=FALSE}
addHTML("primary-metric.html", section = 2, codeChunk = FALSE)
```

### Secondary Metrics {-}

```{r, eval=TRUE, echo=FALSE}
addHTML("secondary-metrics.html", section = 2, codeChunk = FALSE)
```

## Validation {#validation}

We generated our regression model based on the OPTIMAL, STATCOPE, and MACRO studies. To test how well this model performed with new data, we used a fourth study, ECLIPSE.

Details of the study can be found here: [Evaluation of COPD Longitudinally to Identify Predictive Surrogate Endpoints](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2944278/).

#### Criteria {-}

- Age: 40 - 75 years old
- COPD diagnosis
- Does not have asthma
- 10 pack years or more of smoking
- at least 1 COPD exacerbation in the last year
