---
title: "Acute COPD Exacerbation Prediction Tool (ACCEPT)"
site: "bookdown::bookdown_site"
runtime: "shiny"
output:
  bookdown::gitbook:
    lib_dir: "book_assets"
    table_css: FALSE
    css: ["pretty-styles.css", "cayman.css", "python.css"]
    config:
      toc:
        collapse: subsection
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      download: null
      search: yes
      fontsettings:
        theme: white
        family: sans
        size: 2
      sharing:
        facebook: yes
        twitter: yes
        google: no
        linkedin: no
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'google', 'twitter', 'linkedin', 'weibo', 'instapaper']
  bookdown::pdf_book:
    keep_tex: yes
---

# Using the Package in R {#packageinR}

## Installation

If you do not have R installed on your computer, please go to [Appendix 1](#appendix1). Once you have R installed, you can install this package in the R console:

```{r chunk-label, eval = FALSE, echo = TRUE}
install.packages("devtools")
devtools::install_github("resplab/accept")
```

## ACCEPT Prediction Functions

### Sample Data {-}

To get started, there is an R data frame with the package of sample patient data. I have printed columns 1-13 and 14-19 separately because there isn't enough space:

```{r sample-patients, eval = TRUE, echo = TRUE}
library(accept)
samplePatients = accept::samplePatients
print(samplePatients[,1:13])
print(samplePatients[,14:19])
```

### Exacerbation Rate {-}

To get a prediction for exacerbation rate, you will need to pass in a patient vector:

```{r, eval=TRUE, echo=FALSE}
library(htmltools)
rawHTML <- paste(readLines("predictACCEPT-inputs.html"), collapse="\n")
HTML(rawHTML)
```

```{r exacerbation-rate, eval = TRUE, echo = TRUE}
results <- predictACCEPT(samplePatients[1,])
print(t(results))
```

```{r exacerbation-rate-2, eval = TRUE, echo = TRUE}
plotExacerbations(results, type="probability")
```

```{r exacerbation-rate-3, eval = TRUE, echo = TRUE}
plotExacerbations(results, type="rate")
```

### Probability of N Exacerbations (Poisson) {-}

We can also calculate the predicted number of exacerbations in a year:

```{r n-exacerbations-1, eval = TRUE, echo = TRUE}
patientResults = predictACCEPT(samplePatients[1,])
exacerbationsMatrix = predictCountProb(patientResults, n = 10, shortened = TRUE)
print(exacerbationsMatrix)
```

The shortened parameter groups the probabilities from 3-10 exacerbations into one category, "3 or more exacerbations." To see all n exacerbation probabilities:

```{r n-exacerbations-2, eval = TRUE, echo = TRUE}
exacerbationsMatrix = predictCountProb(patientResults, n = 10, shortened = FALSE)
print(exacerbationsMatrix)
```

To visualize the matrix as a heatmap, we can use the function **plotHeatMap**:

```{r n-exacerbations-3, eval = TRUE, echo = TRUE}
plotHeatMap(patientResults, shortened = FALSE)
```
